---
title: "Marketing Spend Analysis"
author: "Phoebe Zhang"
format: 
  html:
   pdf:
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Load required libraries
library(tidyverse)
library(datefixR)
library(vtable)
```

# Introduction
This report investigates if increased personalised marketing campaign spend per customer leads to higher purchase amounts. Understanding this relationship between marketing spend and customer purchase amounts is crucial for evaluating the effectiveness of investment on marketing efforts. The findings from this report can inform future budget allocations around marketing strategies.

# Research Question
This report aims to investigate whether increased marketing spend per customer leads to higher purchase amounts. The relationship between these two variables will be analysed by utilising customer-level data from a personalised marketing campaign, including marketing spend and recorded customer purchase amounts.

# Dataset Introduction
The dataset records customer-level data from a personalised marketing campaign. The data set consists of amount spent on marketing per customer, the number of times the customer visited the platform, the purchase amounts of each customer and each customers sign-up dates. The customers are classified into segments (Consumer, SMB, Enterprise) to reflect different market types and the notes provides information about each customer's marketing interactions.


```{r, echo = FALSE, include = FALSE}
# Read in the raw .csv data file
raw_customer_data <- read.csv("data/37300458_Phoebe Zhang.csv")

# Explore the raw data
head(raw_customer_data)
summary(raw_customer_data)

# Convert all "N/A" and empty null values in the raw data to NA
raw_customer_data[raw_customer_data == "N/A"] <- NA
raw_customer_data[raw_customer_data == ""] <- NA

clean_customer_data <- raw_customer_data %>%
    mutate(
        # Standardise signup_date formats
        signup_date = fix_date_char(signup_date),
        
         # Convert variables to factor to represent categorical data
        segment = as.factor(segment),
        notes = as.factor(notes)
        ) %>%
    # Filter out NA purchase_amounts
    filter(
        !is.na(purchase_amount)
    )
```

## Cleaning Process
The raw data set contained lots of "N/A" and empty null values, these were converted to *NA* to ensure consistency across variables. The raw dataset also contained inconsistent date formats, these were standardised and converted into a consistent date format across all rows. Categorical variables, such as segment and notes, were converted to factor types to properly represent their categorical nature. Rows with missing customer purchase amount data was removed from the data set as unknown purchase amounts would not be useful for this analysis. The cleaning process ensures that the final data set is consistent and complete for meaningful statistical analysis of the relationship between marketing spend and customer purchase amounts.

## Dataset Description
The processed data set contains numeric variables (marketing_spend, visits, purchase_amount), categorical variables (segment, notes), a date time variable (signup_date) and a unique character variable (customer_id). There are `r ncol(clean_customer_data)` variables and `r nrow(clean_customer_data)` observations in the cleaned data set.

```{r}
#| label: tbl-summary-stats
#| tbl-cap: "Summary statistics of numeric variables (marketing spend, visits, purchase amount) in the data set"

# Get summary statistics of the numeric variables
clean_customer_data %>%
    select(marketing_spend, visits, purchase_amount) %>%
    sumtable(title = NA)
```

[@tbl-summary-stats] shows the summary statistics of the numeric variables in the data set. From [@tbl-summary-stats], it appears that the variables, marketing_spend and purchase_amount are highly variable, with ranges spanning from $1.8 to $297 and $0 to $307 respectively, which will allow the analysis to capture the extremes of the relationship between the two variables. The average amount spent on marketing per customer is $148 with most marketing campaigns spending between $74 and $215 and the average customer purchase amount is $135 with most customers spending between $76 and $199. The visits variable shows less variability with a mean of 6.1 and most customers visiting the platform between 4 to 8 times.

[@tbl-summary-stats] provides insights on the distribution of the numeric variables of the data, showing average values, ranges and potential outliers, which can be useful for the interpretation of results later on in the analysis.

```{r}
#| label: fig-custs-per-segment
#| fig-cap: "Number of customers in each different segment (Consumer, Enterprise, SMB) in the data"

# Plot the count of customers in each segment as a bar chart
ggplot(clean_customer_data, aes(x = segment)) +
    geom_bar() +
    labs(x = "Segment",
         y = "Count of Customers",
         title = "Count of Customers by segment") +
    theme_minimal()
```