---
title: "Marketing Spend Analysis"
author: "Phoebe Zhang"
format: 
  html: default
  pdf: default
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Load required libraries
library(tidyverse)
library(vtable)
library(ggplot2)
```

# Introduction
This report investigates if increased personalised marketing campaign spend per customer leads to higher purchase amounts. Understanding this relationship between marketing spend and customer purchase amounts is crucial for evaluating the effectiveness of investment and return on marketing efforts. The findings from this report can inform future budget allocations around marketing strategies.

# Research Question
This report aims to investigate whether increased marketing spend per customer leads to higher purchase amounts. The relationship between these two variables will be analysed by utilising customer-level data from a personalised marketing campaign, including marketing spend and recorded customer purchase amounts.

# Dataset Introduction
The dataset contains records of customer-level data from a personalised marketing campaign. The dataset consists of a unique identifier for each customer, the amount spent on marketing per customer, and the purchase amounts of each customer. The customers are classified into segments (Consumer, SMB, Enterprise) to reflect different market types.

```{r, echo = FALSE, include = FALSE}
# Read in the raw .csv data file
raw_customer_data <- read.csv("data/37300458_Phoebe Zhang.csv")

# Explore the raw data
head(raw_customer_data)
summary(raw_customer_data)

# Convert all "N/A" and empty null values in the raw data to NA
raw_customer_data[raw_customer_data == "N/A"] <- NA
raw_customer_data[raw_customer_data == ""] <- NA

clean_customer_data <- raw_customer_data %>%
    mutate(
        # Convert segment variable to factor to represent categorical data
        segment = as.factor(segment)
    ) %>%
    # Filter out NA purchase_amounts
    filter(!is.na(purchase_amount)) %>%
    # Filter out notes, signup_date and visits columns
    select(-notes, -signup_date, -visits)
```

## Cleaning Process
The raw dataset contained lots of "N/A" and empty null values; these were converted to *NA* to ensure consistency across variables. The categorical variable, `segment`, was converted into a factor type to properly represent its categorical nature. Rows with missing customer purchase amount values were removed from the dataset as unknown purchase amounts would not be useful for this analysis. The `notes` variable was filtered out as it was unstructured and sparse, and is not essential for answering the main research question. The `signup_date` and `visits` variables were also filtered out as they were not necessary for answering the main research question. The cleaning process ensures that the final dataset is consistent and complete for meaningful statistical analysis of the relationship between marketing spend and customer purchase amounts.

## Dataset Description
The processed dataset contains:

* `customer_id`: a unique customer identifier
* `marketing_spend`: a numeric variable representing the amount spent on the marketing for the given customer
* `purchase_amount`: a numeric variable representing the amount spent by the customer
* `segment`: a categorical variable classifying the customer as a Consumer, SMB, or Enterprise

There are `r ncol(clean_customer_data)` variables and `r nrow(clean_customer_data)` observations in the cleaned dataset.
```{r}
#| label: tbl-summary-stats
#| tbl-cap: "Summary statistics of numeric variables (marketing spend, purchase amount) in the dataset"

# Get summary statistics of the numeric variables
clean_customer_data %>%
    select(marketing_spend, purchase_amount) %>%
    sumtable(title = NA)
```

[@tbl-summary-stats] shows the summary statistics of the numeric variables in the dataset. From [@tbl-summary-stats], it appears that the variables, `marketing_spend` and `purchase_amount`, are highly variable, with ranges spanning from $1.8 to $297 and $0 to $307, respectively, which will allow the analysis to capture the extremes of the relationship between the two variables. The average amount spent on marketing per customer is $148, with most marketing campaigns spending between $74 and $215, and the average customer purchase amount is $135, with most customers spending between $76 and $199.

[@tbl-summary-stats] provides insights on the distribution of the numeric variables of the data, showing average values, ranges and potential outliers, which can be useful for the interpretation of results later on in the analysis.

```{r}
#| label: fig-custs-per-segment
#| fig-cap: "Number of customers in each different segment (Consumer, Enterprise, SMB) in the data"

# Plot the count of customers in each segment as a bar chart
ggplot(clean_customer_data, aes(x = segment)) +
    geom_bar() +
    labs(x = "Segment",
         y = "Count of Customers",
         title = "Count of Customers by segment") +
    theme_minimal()
```

[@fig-custs-per-segment] is a bar chart which shows that the observations in this data mostly consists of Consumer customers, where Enterprise customers contribute the least to this data.

[@fig-custs-per-segment] shows how the customers are distributed across the different segments (Consumer, SMB, Enterprise), and provides background on which segments contribute most to this data. This can prove useful later on if the marketing team would like to analyse marketing spend based on each different customer segment.

# Results
```{r}
#| label: fig-market-cust-plot
#| fig-cap: "Marketing spend per customer, against each customer's purchase amount in dollars ($)"

ggplot(clean_customer_data, aes(x = marketing_spend, y = purchase_amount, color = segment)) +
    geom_point() +
    labs(x = "Marketing Spend ($)", 
         y = "Customer Purchase Amount ($)",
         title = "Marketing Spend against Customer Purchase Amount" ) +
    theme_minimal()
```

[@fig-market-cust-plot] shows that there is a strong positive relationship between marketing spend and customer purchase amount. As marketing spend increases, customer purchase amount also tends to increase. There is some variability in the data shown by the dispersion of the points on the plot, but the overall trend is positive. At lower marketing spend on personalised campaigns, most customers make no purchase, or the purchase amount is small. At higher amounts of money spent on personalised marketing campaigns, customer purchases tend to be generally higher. This suggests that higher marketing spend per customer does lead to higher purchase amounts. 

Customers with marketing spend amounts below $50 had an average purchase amount of $`r round(mean(clean_customer_data$purchase_amount[clean_customer_data$purchase_amount < 50]), 2)`, whereas customers with marketing spend amounts above $250 had an average purchase amount of $`r round(mean(clean_customer_data$purchase_amount[clean_customer_data$purchase_amount > 250]), 2)`, showing that there is a general upward trend between the two variables. The correlation between marketing spend and customer purchase amount is `r round(cor(clean_customer_data$marketing_spend, clean_customer_data$purchase_amount), 2)`. This correlation coefficient is positive and close to 1, indicating a strong positive relationship between the two variables. These statistics further support the conclusion that higher marketing spend per customer leads to higher purchase amounts.