---
title: "Marketing Spend Analysis"
author: "Phoebe Zhang"
format: 
  html: default
  pdf: default
geometry: margin = 1in
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Load required libraries
library(tidyverse)
library(datefixR)
library(knitr)
library(ggplot2)
```

# Introduction
This report investigates if increased personalised marketing campaign spend per customer leads to higher purchase amounts. Understanding this relationship between marketing spend and customer purchase amounts is crucial for evaluating the effectiveness of investment and return on marketing efforts. The findings from this report can inform future budget allocations around marketing strategies.

# Research Question
This report aims to investigate whether increased marketing spend per customer leads to higher purchase amounts. The relationship between these two variables will be analysed by utilising customer-level data from a personalised marketing campaign, including marketing spend and recorded customer purchase amounts.

# Dataset Introduction
The dataset contains records of customer-level data from a personalised marketing campaign, showing how much was spent on marketing for each customer (`marketing_spend`) and the customer's corresponding purchase amount (`purchase_amount`). The dataset also contains information on the customer's sign-up date, id, number of visits, segment, and associated notes.

```{r, echo = FALSE, include = FALSE}
# Read in the raw .csv data file
raw_customer_data <- read_csv("data/37300458_Phoebe Zhang.csv")

# Explore the raw data
head(raw_customer_data)
summary(raw_customer_data)

# Convert all "N/A" and empty null values in the raw data to NA
raw_customer_data[raw_customer_data == "N/A"] <- NA
raw_customer_data[raw_customer_data == ""] <- NA

clean_customer_data <- raw_customer_data %>%
    mutate(
        # Standardise signup_date formats
        signup_date = fix_date_char(signup_date),
        
        # Convert variables to factor to represent categorical data
        segment = as.factor(segment),
        notes = as.factor(notes)
    ) %>%
    # Filter out rows with missing purchase_amounts and marketing_spend
    filter(
        !is.na(purchase_amount),
        !is.na(marketing_spend)
    )
```

## Cleaning Process
The raw dataset contained lots of N/A and empty NULL values; these were converted to *NA* to ensure consistency across variables. Inconsistent date formats were standardised into a consistent date format across all rows. Categorical variables, such as `segment` and `notes`, were converted to factor types to properly represent their categorical nature. Rows with missing marketing spend and customer purchase amount values were removed as these unknown values would not be useful for the analysis. The cleaning process ensures that the final dataset is consistent and complete for meaningful statistical analysis of the relationship between marketing spend and customer purchase amounts.

## Dataset Description
The processed dataset contains:

* `customer_id`: a unique customer identifier
* `marketing_spend`: a numeric variable representing the amount spent on the marketing for the given customer
* `visit`: a numeric variable representing the number of times the given customer visited the platform
* `purchase_amount`: a numeric variable representing the amount spent by the customer
* `signup_date`: a date variable representing the date the customer signed up to the platform
* `segment`: a categorical variable classifying the customer as a Consumer, SMB, or Enterprise
* `notes`: a categorical variable representing customer response to marketing outreach

There are `r ncol(clean_customer_data)` variables and `r nrow(clean_customer_data)` observations in the cleaned dataset.
```{r}
#| label: tbl-average-stats
#| tbl-cap: "Averages of numeric variables (marketing spend, visits, purchase amount) in each customer segment (Consumer, Enterprise, SMB)"

# Get averages of numeric variables grouped by customer segments
clean_customer_data %>%
    group_by(segment) %>%
    summarise(
        customer_count = n(),
        avg_marketing_spend = mean(marketing_spend),
        avg_purchase_amount = mean(purchase_amount),
        avg_visits = mean(visits)
    ) %>%
    kable(digits = 2)
```

[@tbl-average-stats] provides a summary of customer characteristics across different segments, offering insight on how marketing spend and customer purchase behaviour varies across different customer segments. From [@tbl-average-stats], it appears that Enterprise customers had the highest average platform visits and return for marketing spend, with an average marketing spend of $109.13 and an average purchase amount of $106.38.

```{r}
#| label: fig-custs-per-segment
#| fig-cap: "Number of customers in each different segment (Consumer, Enterprise, SMB)"

# Plot the count of customers in each segment as a bar chart
ggplot(clean_customer_data, aes(x = segment, fill = segment)) +
    geom_bar() +
    labs(
        x = "Segment",
        y = "Number of Customers",
        title = "Count of Customers by Segment"
    ) +
    theme_minimal()
```

[@fig-custs-per-segment] shows how the customers are distributed across segments (Consumer, SMB, Enterprise), and provides background on which segments contribute most to the dataset. From [@fig-custs-per-segment], it appears that the observations in the dataset mostly consist of Consumer customers, and Enterprise customers contribute the least records to the dataset. This can prove useful later on if the marketing team would like to analyse marketing spend based on each different customer segment. 

# Results
```{r}
#| label: fig-market-cust-plot
#| fig-cap: "Marketing spend per customer, against each customer's purchase amount in dollars"

ggplot(clean_customer_data, aes(x = marketing_spend, y = purchase_amount)) +
    geom_point() +
    labs(
        x = "Marketing Spend ($)", 
        y = "Customer Purchase Amount ($)",
        title = "Marketing Spend against Customer Purchase Amount"
    ) +
    theme_minimal()
```

[@fig-market-cust-plot] shows that there is a strong positive relationship between marketing spend and customer purchase amount. As marketing spend increases, customer purchase amount also tends to increase. There is some variability in the data shown by the dispersion of the points on the plot, but the overall trend is positive. At lower marketing spend on personalised campaigns, most customers make no purchase, or the purchase amount is small. At higher amounts spent on personalised marketing campaigns, customer purchases tend to be generally higher. This suggests that higher marketing spend per customer does lead to higher purchase amounts. 

Customers with marketing spend below $50 had an average purchase amount of $`r round(mean(clean_customer_data$purchase_amount[clean_customer_data$purchase_amount < 50]), 2)`, whereas customers with marketing spend above $250 had an average purchase amount of $`r round(mean(clean_customer_data$purchase_amount[clean_customer_data$purchase_amount > 250]), 2)`, showing that there is a general upward trend between the two variables. The correlation between marketing spend and customer purchase amount is `r round(cor(clean_customer_data$marketing_spend, clean_customer_data$purchase_amount), 2)`. This correlation coefficient is positive and close to 1, indicating a strong positive relationship between the two variables. These statistics further support the conclusion that higher marketing spend per customer leads to higher purchase amounts.

```{r, message=FALSE}
#| label: fig-by-segment-plot
#| fig-cap: "Marketing spend against purchase amount for each customer in dollars ($) by segment"

# Plot marketing spend against purchase amount by segment
ggplot(
    clean_customer_data, 
    aes(
        x = marketing_spend, 
        y = purchase_amount, 
        color = segment
    )
) + 
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    facet_wrap(~segment) +
    theme_minimal()
```

[@fig-by-segment-plot] displays marketing spend against purchase amount for each customer by segment (Consumer, Enterprise, SMB). In all 3 segments, the relationship between marketing spend and purchase amount is positive, suggesting that higher marketing spend generally leads to higher customer purchase amounts across all 3 segments. The spread and slope of the data varies across the 3 segments, with Consumer and SMB customers capturing a wider range of customer purchase amounts at higher marketing spend levels, while Enterprise customers captures a much smaller range due to fewer observations. The plot indicates that increased marketing spend tends to lead to higher purchase amounts, regardless of customer segments.